Modern BTree\\ techniques

### 2020-01-15 btree/presentation.tex
hybrid indexes

\usepackage[utf8]{inputenc}
\usepackage[russian, english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{mindmap}
\usetikzlibrary{shapes.misc, fit}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usetikzlibrary{trees}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{csquotes}
\usetikzlibrary{tikzmark}

@@

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}

@@

@@

            \node[draw, inner sep=0.2cm, behind path, dashed,
                fit=(node1)(node2)(node2-extra)
            ] (extent-1) {};

            \node[draw, inner sep=0.2cm, behind path, dashed,
                fit=(node3)(node4)(node5)(node5-extra)
            ] (extent-2) {};
        \end{tikzpicture}

@@

            \node[draw, inner sep=0.2cm, behind path, dashed,
                fit=(node1)(node2)(node2-extra)
            ] (extent-1) {};

            \node[draw, inner sep=0.2cm, behind path, dashed,
                fit=(node3)(node4)(node5)(node5-extra)
            ] (extent-2) {};

@@

            \node[arrow-pointer, right=2cm of node9.center, rotate=-135, xshift=-2.7cm, yshift=-1cm] {};
            \node[draw, behind path, fit=(node4)(node9)(node10), rounded corners, dashed] {};
            \node[below=3cm of node3.center, minimum height=1cm] {};

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[russian, english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{mindmap}
\usetikzlibrary{shapes.misc, fit}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usetikzlibrary{trees}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{csquotes}
\usetikzlibrary{tikzmark}
\usetikzlibrary{arrows.meta}

@@

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}

@@

\newcommand{\smallbtree}[6] {
    \node[smallbtree-node, #4, #3] (node#11) {};

    \node[smallbtree-node, #4, below=0.4cm * #2 of node#11.south west, xshift=-0.4cm * #2] (node#12) {};
    \node[smallbtree-node, #4, below=0.4cm * #2 of node#11.south] (node#13) {};
    \node[smallbtree-node, #4, below=0.4cm * #2 of node#11.south east, xshift=0.4cm * #2] (node#14) {};

    \draw[->, btree-line, #5] (node#11) -- (node#12);
    \draw[->, btree-line, #5] (node#11) -- (node#13);
    \draw[->, btree-line, #5] (node#11) -- (node#14);
    \node[#6, rounded corners=0.2cm, inner sep=0.2cm,
          fit=(node#11)(node#12)(node#13)(node#14)] (nodebox#1) {};

    % redraw
    \node[smallbtree-node, #4, #3] (node#11) {};

    \node[smallbtree-node, #4, below=0.4cm * #2 of node#11.south west, xshift=-0.4cm * #2] (node#12) {};
    \node[smallbtree-node, #4, below=0.4cm * #2 of node#11.south] (node#13) {};
    \node[smallbtree-node, #4, below=0.4cm * #2 of node#11.south east, xshift=0.4cm * #2] (node#14) {};

    \draw[->, btree-line, #5] (node#11) -- (node#12);
    \draw[->, btree-line, #5] (node#11) -- (node#13);
    \draw[->, btree-line, #5] (node#11) -- (node#14);
}

@@

            \node[box, inner sep=0.2cm,
                fit=(node1)(node2)(node2-extra)
            ] (extent-1) {};

            \node[box, inner sep=0.2cm,
                fit=(node3)(node4)(node5)(node5-extra)
            ] (extent-2) {};
        \end{tikzpicture}

@@

            \node[box, inner sep=0.2cm,
                fit=(node1)(node2)(node2-extra)
            ] (extent-1) {};

            \node[box, inner sep=0.2cm,
                fit=(node3)(node4)(node5)(node5-extra)
            ] (extent-2) {};

@@

            \node[arrow-pointer, right=2cm of node9.center, rotate=-135, xshift=-2.7cm, yshift=-1cm] {};
            \node[box, fit=(node4)(node9)(node10)] {};
            \node[below=3cm of node3.center, minimum height=1cm] {};

@@

    \begin{center}
    \textbf{Hybrid indexes}
    \vspace{1cm}

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \smallbtree{1}{2.5}{}{}{}{green-box}
            \node[draw, color=white, left=4cm of node11, minimum width=1.2cm,
                  minimum height=0.4cm, yshift=-0.6cm] (node21) {};
            \node[bloom-filter, color=white, above=0.5cm of node21] (bloom) {};
            \node[arrow-pointer, left=0.5cm of node21.west, color=white] {\footnotesize{Insert}};
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \smallbtree{1}{2.5}{}{}{}{green-box}
            \smallbtree{2}{1}{left=4cm of node11, yshift=-0.6cm}{color=blue!20}{color=blue!20}{blue-box}
            \node[bloom-filter, color=white, above=0.5cm of node21] (bloom) {};
            \node[arrow-pointer, left=0.5cm of nodebox2.west, color=white] {\footnotesize{Insert}};
        \end{tikzpicture}

        \onslide<3>
        \begin{tikzpicture}[]
            \smallbtree{1}{2.5}{}{}{}{green-box}
            \smallbtree{2}{1}{left=4cm of node11, yshift=-0.6cm}{}{}{blue-box}
            \node[bloom-filter, color=white, above=0.5cm of node21] (bloom) {};
            \node[arrow-pointer, left=0.5cm of nodebox2.west, color=white] {\footnotesize{Insert}};
        \end{tikzpicture}

        \onslide<4>
        \begin{tikzpicture}[]
            \smallbtree{1}{2.5}{}{}{}{green-box}
            \smallbtree{2}{1}{left=4cm of node11, yshift=-0.6cm}{}{}{blue-box}
            \node[bloom-filter, above=0.5cm of node21, fill=redBad] (bloom) {};
            \node[arrow-pointer, left=0.5cm of nodebox2.west, color=white] {\footnotesize{Insert}};
        \end{tikzpicture}

        \onslide<5>
        \begin{tikzpicture}[]
            \smallbtree{1}{2.5}{}{}{}{green-box}
            \smallbtree{2}{1}{left=4cm of node11, yshift=-0.6cm}{}{}{blue-box}
            \node[bloom-filter, above=0.5cm of node21, fill=redBad] (bloom) {};
            \node[arrow-pointer, left=0.5cm of nodebox2.west] {\footnotesize{Insert}};
        \end{tikzpicture}

        \onslide<6>
        \begin{tikzpicture}[]
            \smallbtree{1}{2.5}{}{}{}{green-box}
            \smallbtree{2}{1}{left=4cm of node11, yshift=-0.6cm}{}{}{blue-box}
            \node[bloom-filter, above=0.5cm of node21, fill=redBad] (bloom) {};
            \node[arrow-pointer, left=0.5cm of nodebox2.west] (inserts) {\footnotesize{Insert}};
            \node[arrow-pointer, above=1.2cm of inserts] {\footnotesize{Read}};
            \draw[->, btree-line] (bloom.south) -- (nodebox2.north);
        \end{tikzpicture}

    \end{overprint}
    \end{center}
\end{frame}
\note{
    Designed for B-Tree, Masstree, Skip List, ART
}
